---
title: "Notebook 05"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. You may also have to hit the broom in
the upper right-hand corner of the window. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
source("../funs/funs_custom.R")
```

I have set the options `message=FALSE` and `echo=FALSE` to avoid cluttering
your solutions with all the output from this code.

## Reading the Data

Today we are going to look at a dataset of short texts taken from a set of 3
American authors:

```{r, message = FALSE}
docs <- read_csv("../data/stylo_us.csv")
anno <- read_csv("../data/stylo_us_token.csv.gz")
```

The prediction task is to determine the identity of the author based on the
text.

## Questions

### Authorship Detection with the Elastic Net

To start, build an elastic net model using the default term-frequency features
to predict the authorship of each text:

```{r, question-01}

```

Now, what is the overall error rate of this model?

```{r, question-02}

```

In the code block below, look at the model coefficients. You may want to limit
the number of results by setting `lambda_num` to something around 30

```{r, question-03}

```

You should notice several interesting things about the model. What are the main
features that are being used here? Do you see any patterns about where the
strongest coefficients are concentrated?

Build another elastic net model but remove the proper nouns from the texts.
Note, we can use `!=` to signal not equal to in R.

```{r, question-04}

```

What is the error rate of the model now? Note how it changes from above.

```{r, question-05}

```

In the code block below, look at the model coefficients. You may want to limit
the number of results by setting `lambda_num` to something around 40.

```{r, question-06}

```

You should now notice that the types of words being used are very different.
What are the primary qualities of (most of) the terms now being selected?

Now, build an elastic net model using the "xpos" column to build frequencies.
These are a more granular way of describing parts of speech.

```{r, question-07}

```

Compute the error rate, comparing it to the previous error rates.

```{r, question-08}

```

We can do a bit better using n-grams of xpos tags. Create a model using 3-grams
of xpos tags.

```{r, question-09}

```

Finally, compute the error rate of this model:

```{r, question-10}

```

How does it compare the word-based model above? It will likely be very close,
though perhaps still slightly worse.

### Authorship Detection with the k-nearest neighbors

Now, let's use the k-nearest neighbors method to build a model. Start by
creating a knn model with all of the default features and k = 1 (the default).

```{r, question-11}

```

Now, print out the error rate of the model. Compare to the same model using
the elastic net.

```{r, question-12}

```

Now, fit a knn model using the 3-gram xpos features and k = 1.

```{r, question-13}

```

Compute the error rate and compare the results using the elastic net.

```{r, question-14}

```

Finally, fit a model using 3-gram xpos features and k = 10.

```{r, question-15}

```

Compute the error rate and compare the previous results. Does setting k equal
to 10 change the model?

```{r, question-16}

```

You should have noticed that knn is significantly worse than the elastic net
model. Take a few moments and think about what the difference between these
models is and why knn is not particularly suitable to the prediction task here.
