---
title: "Notebook 06"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. You may also have to hit the broom in
the upper right-hand corner of the window. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
source("../funs/funs_custom.R")
```

I have set the options `message=FALSE` and `echo=FALSE` to avoid cluttering
your solutions with all the output from this code.

## Reading the Data

Today we are going to look at a dataset of short texts taken from a set of 5
British authors:

```{r, message = FALSE}
docs <- read_csv("../data/stylo_uk.csv")
anno <- read_csv("../data/stylo_uk_token.csv.gz")
```

The prediction task is to determine the identity of the author based on the
text.

## Questions

### Baseline Elastic Net

Start by fitting an elastic net model with all of the default parameters. We
will need this several different times, so save the model with a unique name
such as `model_enet`:

```{r, question-01}

```

Compute the error rate of the elastic net model on the training and validation
sets.

```{r, question-02}

```

### Baseline Gradient Boosted Trees

Now, fit a gradient boosted tree model using the default parameters:

```{r, question-03}

```

Compute the error rate of this model:

```{r, question-04}

```

Note that this model is not particularly good. The problem is that we need
a significantly larger set of trees. Create a new gradient boosted tree model
using 1000 trees:

```{r, question-05}

```

Compute the error rate now for the training and validation sets. Do this below
the normal way, but note that they should also be given in the output printed
above.

```{r, question-06}

```

How does the error rate compare to the elastic net model? Make sure to look at
both the training and validation results.

### Variable Importance

Now, look at the coefficients for the elastic net model. There are a lot, so
you may want to limit the lambda number to something around 30.

```{r, question-07}

```

And, for comparison, look at the importance scores for the gradient boosted
trees:

```{r, question-08}

```

How do the two lists compare to one another? Is there a lot of overlap? Is there
any particular pattern to the differences?

### Data Visualization

For the first visualization in today's notebook, take the elastic net model
you built above (`model_net`) and select the attached `docs` table. Filter to
just include the validation set and arrange the data from the highest value of
`pred_value` (the probability that the model thinks it is correct). Then, slice
to include just the first 1000 rows and compute the error rate. Note that this
should be much lower than the rate you had above. Play around with the number 
of rows that are included in the summary and try to understand the pattern.

```{r, question-09}

```

Now, we want to create a plot that visualizes the number of points included in
the summary. To do this, replace the summary command you have above with a 
mutate verb, using the function `cummean` in place of `mean`. Also, create a
variable with the command `prop = row_number() / n()`. Take the result and 
plot the proportion of the corpus predicted on the x-axis and the error rate on
the y-axis using a line geometry.

```{r, question-10}

```

Now, take the predictions from the top 4500 most confident predictions in the
validation data (it's about 75% of the values) and build a confusion matrix.
Take a moment to understand what it's telling us, keeping in mind that the
data are balanced between the labels.

```{r, question-11}

```

Finally, repeat the last question on Notebook 5 using just the validation data
from the dataset here. Order the authors from the highest proportion of verbs
used to the lowest. 

```{r, question-12}

```

Using the intersection of the confidence intervals (it's a rough measurement,
but a reasonably accurate measurement of statistical significant), you should
see that the authors fall into three rough buckets.
