---
title: "Notebook 04"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. You may also have to hit the broom in
the upper right-hand corner of the window. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
source("../funs/funs_custom.R")
options(dsst.traceit = FALSE)
```

I have set the options `message=FALSE` and `echo=FALSE` to avoid cluttering
your solutions with all the output from this code.

## Reading the Data

To start, let's read in some text data to study for today. The dataset for this
notebook consists of a set of movie reviews.

```{r, message = FALSE}
docs <- read_csv("../data/imdb_polarity_mod.csv.gz")
anno <- read_csv("../data/imdb_polarity_mod_token.csv.gz")
```

The label we are trying to predict is whether a review is a 4 (out of 10) or a
7 (out of 10).

## Questions

### Model with defaults

Start by building an elastic net model for this data using the default values.
Note that you will have to set the parameter `label_var` equal to "category"
because the label has a different name.

```{r, question-01}

```

Now compute the error rate of the model.

```{r, question-02}

```

Then, the model coefficients.

```{r, question-03}

```

And finally, use the following code to see the cross-validation curve.

```{r, question-04}

```

Now, before proceeding, take a moment to look at all of the results and
understand what the model is telling you about the data and how well we can
make predictions of the label.

### Model using verbs

Now, fit an elastic net model using only those lemmas labeled as a VERB.

```{r, question-05}

```

Now compute the error rate of the model.

```{r, question-06}

```

Then, the model coefficients.

```{r, question-07}

```

And finally, plot the cross-validation curve.

```{r, question-08}

```

Take a moment to compare the results to the previous model. How well does this
do compared to all of the parts of speech? Can you make any sense of the
coefficients in a new way that tells you something about the data? How about
the cross-validation curve, does it tell you anything about the model?

### Model using ngrams

Now, fit an elastic net model using bigrams and unigrams of all the lemmas.

```{r, question-09}

```

Now compute the error rate of the model.

```{r, question-10}

```

Then, the model coefficients.

```{r, question-11}

```

And finally, plot the cross-validation curve.

```{r, question-12}

```

How does this model do compared to the original model? How many of the top
variables are bigrams compared with unigrams? How much new do you learn about
the data from the bigram model?

### Model using pos values

Finally, fit an elastic net model using just the upos codes.

```{r, question-13}

```

Now compute the error rate of the model.

```{r, question-14}

```

Then, the model coefficients.

```{r, question-15}

```

And finally, plot the cross-validation curve.

```{r, question-16}

```

How does this model do compared to the original one? Probably not great, but
notice that it does do better than random guessing. What can be learned from
the model? Can you understand why any of the coefficients have the signs that
they do? Does this help you understand any results above?

### Best model?

If you have some extra time, try to find the most predictive model that you
can create in the block below.

```{r, question-17}

```
