---
title: "Notebook 10"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. You may also have to hit the broom in
the upper right-hand corner of the window. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
source("../funs/funs_custom.R")
```

I have set the options `message=FALSE` and `echo=FALSE` to avoid cluttering
your solutions with all the output from this code.

## Reading the Data

Today we will again look at a subset of a well-known text analysis corpus
call NewsGroups-20.

```{r, message = FALSE}
docs <- read_csv("../data/newsgroups.csv.bz2")
anno <- read_csv("../data/newsgroups_token.csv.bz2")
```

## Questions

### Clustering

Use K-means to cluster the categories with K equal to 5 (that's the default).
Look at the clusters and see if they match the plots you made in the previous
notebook.

```{r, question-01}

```

Now, run hierarchical clustering on the categories. Save the model as an R
object.

```{r, question-02}

```

Plot the hierarchical clustering and notice what additional information it
gives compared to K-means.

```{r, question-03}

```

Cut the tree into 5 clusters and compare to the K-means clusters.

```{r, question-04}

```

### Words

Produce a PCA plot of the nouns, using the option invert = TRUE. You will want
to adjust the parameter `min_df` to be `0.01`.

```{r, question-05}

```

Produce a PCA plot of the verbs, using the option invert = TRUE. You will want
to adjust the parameter `min_df` to be `0.01`.

```{r, question-06}

```

What interesting patterns do you see in the previous two plots?
