---
title: "Title Here"
output:
  html_document
---

This is a template RMarkdown document that you can use for your own
work. Unlike the notebooks, it puts all of the code that you need
directly in  the code, so you can place this file anywhere on your
machine (as well as create as many copies as you need) and it should
still work. Note that you can delete/replace this paragraph of text
as well as the following.

To start, you should always have a block of code that does all of
the setup and options that you need for the rest of the document.
As with the notebooks, you need to rerun this each time you restart
R. You can add more library commands if needed or delete ones you 
do not need or prefix them with a pound sign (as I've done with the
dsst package).

```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(jsonlite)
library(ggrepel)
library(stringi)
# library(dsst) 

options(dplyr.summarise.inform = FALSE)
options(ggrepel.max.overlaps = Inf)
```

Now, you can add whatever text or code you would like below. The
part above should give you the the setup we had in class.

```{r}

```