---
title: "Project 01 -- Instructions"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

```{r, include=FALSE}
```

**Due Date**: 01 February 2023

## General Instructions

This page outlines the instructions for the first project. You should
have a file `project01.Rmd` in your RStudio workspace where you can
work on the project. I find that students prefer having a consistent
format for the projects, so I will attempt to keep the format the same
throughout the semester.

On the due-date, your group is responsible for completing two elements:

1. Slides presenting your results, also submitted in your Box folder.
2. A presentation based on your slide show. This should be 5 minutes in length.

As described on the syllabus, the project will be graded using a rubric, which
can be found [here](../extra/project_rubric.pdf).

## Instructions

This project uses a collection of book reviews from the website Good Reads.
The data have been taken from the following paper:

> Mengting Wan, Julian McAuley, "Item Recommendation on Monotonic Behavior Chains",
> in RecSys'18.

Groups will be assigned different genres to look at, so you will have a slightly
different analysis compared to the rest of the class. You can read more about
the collection as a whole [here](https://sites.google.com/eng.ucsd.edu/ucsdbookgraph/home#h.p_VCP_qovwtnn1).

There are two different prediction tasks that you may look at. The first is to
classify the score of the book on a scale from 1-5. The second is to predict
the gender of the author of the book being described. Here are some things you
might consider addressing:

- How accurately can you classify the groups based on the text on the
validation data? What kinds of errors does the model tend to make?
- What terms most distinguish the groups? If you look at just verbs or
adjectives does anything different pop up?
- Looking at negative example from your best model, do you find it easy to
identify the correct class or not?
- If any of the most indicative terms seem surprising, apply KWiC to the
data. Does that help explain why the term is predictive?
- Is there a difference in the parts of speech used between the groups?
- Create a confusion matrix and describe any patterns (interesting or not).

At the end of the day, though, your task is an open-ended one. I want you to
explore the data using the techniques we have learned so far and then produce
a short presentation showing your results to the rest of the class.

## Presentation

Your presentation should be in the form of a set of slides. You can build these
in whatever software you would like, but please create a PDF version to submit
on Box by the project deadline.

The exact format of the slides is up to you, however there should be a final
slide titled "Synthesis" that summarises what you have learned about the data
from your analysis. It should take a big picture view of the analysis and not
be overly focused on the models.

You will find that most (perhaps all) of the results you want to show are
tables and example reviews. Do not use screen shots for these! Screen shots are
messy and not ideal. Instead, I suggest using the function `dsst_clipboard()`
to copy information to the clipboard. This can then be pasted into a spreadsheet
program or (in some cases) directly into a presentation. For example:

```{r, eval=FALSE}
dsst_neg_examples(model) %>% dsst_clipboard()
```

For visualisations, you should use the function `ggsave` to store the
most recent plot as a JPG or PNG file.

## Groups

Groups are assigned different genres to look at in their project. You
should be able to download your data set from within the `project01.Rmd` file.

**Group 1**: `children_b`

**Group 2**: `comics_graphic_b`

**Group 3**: `fantasy_paranormal_b`

**Group 4/9**: `history_biography_b`

**Group 5**: `mystery_thriller_crime_b`

**Group 6**: `poetry_b`

**Group 7**: `romance_b`

**Group 8**: `young_adult_b`

## Notes

While working through the project, I typically find that many groups ask for
help writing the same bits of code. Any notes that I want to share about how
to do specific tasks will be added here as we work through the project.
