---
title: "11. Topic Models"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
options(dsst.traceit = FALSE)
```

*As mentioned in class, you do not need to read the notes for today ahead of
time. We will discuss them together in class. Please spend the extra time 
working on your Project 3.*

## Load Wikipedia Data

We will start by looking at a collection of Wikipedia pages related to sovereign
states (i.e., countries):

```{r, message=FALSE}
docs <- read_csv(file.path("..", "data", "wiki_list_of_sovereign_states.csv"))
anno <- read_csv(file.path("..", "data", "wiki_list_of_sovereign_states_anno.csv.gz"))
```

Today we will use these to illustrate a new technique for understanding a large
collection of documents. 

## Latent Dirchlet Allocation: Method

Now, we will investigate a method for *topic modeling*. This is an
unsupervised task that seeks to identify topics within a corpus of text. What
exactly is a topic? Mathematically speaking, it is usually defined as a
probability distribution over a collection of words. Words that have a high
probability within a topic tend to classify the topics themes in a colloquial
sense. For example, a topic that captures the idea of baseball would have high
probabilities on words such as "base", "player", "strike", "umpire", "team",
and so forth.

We will be use a model today called Latent Dirchlet Allocation, or more commonly
LDA. Given a fixed number of topics and fixed set of words (called a *lexicon*),
LDA assumes that documents consist of a random collection of words
constructed according to the following model:

1. Each document is randomly partitioned into topics. For example, one document
may be 20% in Topic A, 70% in Topic B, and 1% in the remaining 10 topics.
2. Each topic is similarly assigned as a probability distribution of all the
available words.
3. When choosing words to create a document, pick a topic at random
proportional to the topic distribution of the document, and then pick a word
proportional to the chosen topic.
4. The number of words in each document is assumed to be fixed, and there is
assumed to be no relationship between the words in each document.

This model is a great example of the adage that "all models are wrong, but
some are useful". Clearly, this is not how documents are constructed, and
words are not independent of one another. However, the approximation is close
enough to produce a useful abstraction for detecting themes within a corpus of
textual documents.

You will notice that the description above is in some ways backwards from
reality. It assumes that we know the distribution of the topics over the words
and documents but do not know what words are present in the documents. In fact,
we know the words but not the topics! This is an example of a Bayesian model.
If we wrote down the assumptions rigorously, we could invert the probabilities
using Bayes' Theorem. That is, instead of knowing the probability of the
documents given the topics, we can determine the probability of the topics given
the documents. It is not possible to do this analytically, however, and a
simulation method is needed to figure out what distribution of topics over the
words and documents is most likely to have produced the observed data.

As a final note about the method, the name comes from the standard distribution
used to determine the topics (the Dirichlet distribution) and the fact that the
topics themselves are never observed (that is, the are *latent*).

## Latent Dirchlet Allocation: Application

Now, let's actually compute an LDA topic model using the function
`dsst_lda_build`. It requires setting the number of topics; 16 is a good
starting number. I tend to use only nouns, adjectives, adverbs, and verbs for
the LDA model. The model can take a few minutes to finish; it should print
out results as the algorithm proceeds.

```{r}
model <- anno %>%
  filter(upos %in% c("NOUN", "VERBS", "ADJ", "ADV")) %>%
  dsst_lda_build(num_topics = 16)
```

The results in the object `model` are structured as data tables. You can use
them directly if you want to do some direct EDA on the object. For example,
here are the documents most associated with each topic:

```{r}
model$docs %>%
  group_by(topic) %>%
  arrange(desc(prob)) %>%
  slice_head(n = 5) %>%
  summarise(countries = paste(doc_id, collapse = "; "))
```

Likewise, here are the words most associated with each topic:

```{r}
model$terms %>%
  group_by(topic) %>%
  arrange(desc(beta)) %>%
  slice_head(n = 5) %>%
  summarise(words = paste(token, collapse = "; "))
```

Otherwise, we return again to the idea of building an interactive visualisation
using JavaScript. We can write all of the model data as a local JSON file with
the `dsst_json_lda` function:

```{r}
dsst_json_lda(model, docs)
```

The results are, by default, stored in the output directory of your class
notes as a file named "". If we go to the website
[Topic Model Visualizer](https://statsmaths.github.io/topic-explo/build/)
and upload the JSON file that we just produced, it will create an interactive
visualisation of the topics for us to explore.
