---
title: "10. Exploratory Analysis"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
options(dsst.traceit = FALSE)
```

## Load the Data

For today's notes, I want to illustrate ways that we can work with the kinds of
data you are looking at with the projects. Specifically, I will look at the
Yelp reviews from Montréal (these have not been assigned to any group):

```{r, message=FALSE}
docs <- read_csv("../data/revs_montreal.csv.bz2")
anno <- read_csv("../data/revs_montreal_token.csv.bz2")
```

I'll start by looking at a PCA plot of the reviews along with a default number
of clusters:

```{r}
anno %>%
  inner_join(select(docs, -text), by = "doc_id") %>%
  dsst_pca(doc_var = "user_name", n_dims = 2) %>%
  dsst_kmeans(n_clusters = 5) %>%
  ggplot(aes(v1, v2, color = factor(cluster))) +
    geom_point() +
    theme_void()
```

Notice that the data are very cluster into two distinct groups. Let's redo
the analysis with just two clusters. I will be using the results a few different
ways, so I save versions of the docs and anno tables here.

```{r}
docs_cluster <- anno %>%
  inner_join(select(docs, -text), by = "doc_id") %>%
  dsst_pca(doc_var = "user_name", n_dims = 2) %>%
  dsst_kmeans(n_clusters = 2) %>%
  mutate(train_id = "train") %>%
  mutate(doc_id = user_name)

anno_cluster <- anno %>%
  inner_join(select(docs, doc_id, user_name), by = "doc_id") %>%
  inner_join(docs_cluster, by = "user_name", suffix = c("_orig", ""))
```

Plotting the results of the clustering shows the clear distinction between the
two groups:

```{r}
docs_cluster %>%
  ggplot(aes(v1, v2, color = factor(cluster))) +
    geom_point() +
    theme_void()
```

We can use the `dsst_metrics` function to find the terms most associated with
each cluster:

```{r}
dsst_metrics(anno_cluster, docs_cluster, label_var = "cluster") %>%
  filter(count > expected) %>%
  group_by(label) %>%
  slice_head(n = 25) %>%
  summarize(token = paste0(token, collapse = " ")) %>%
  getElement("token")
```

As you might have guessed, the clusters are related to the language of the 
review, either English or French. Now, let's try to label the language of a
review. We could use a fancy model, but I think it's interesting to try to do
this with just the data manipulation tools we have access to. To do this, I
will save the top-25 words for each cluster:

```{r}
lang_df <- dsst_metrics(anno_cluster, docs_cluster, label_var = "cluster") %>%
  filter(count > expected) %>%
  group_by(label) %>%
  slice_head(n = 25) %>%
  ungroup() %>%
  mutate(lang = if_else(label == 1, "English", "Français")) %>%
  select(lemma = token, lang)

lang_df
```

Now, we can do an inner join to the annotation data and count how many terms 
from each of these two sets constitute each document. Here, we visualize 
the proportion of each document that comes from one of the two sets:

```{r}
anno %>%
  inner_join(lang_df, by = "lemma") %>%
  group_by(doc_id) %>%
  summarize(percent_fr = mean(lang == "Français")) %>%
  ggplot(aes(percent_fr)) +
    geom_histogram(bins = 50, color = "black", fill = "white")
```

We see that most documents are mostly in either English or French, though many
have a few words from the other set. I looked at these, and they are mostly 
either the names of businesses or (mostly in French), short English expressions
inserted into the text.

There are a small number of reviews that are in the middle of our plot above 
(it might be hard to see in the compiled notes, but I could tell in RStudio).
Let's take a look at a few of these:

```{r}
set.seed(1L) # to make sure I always take the same set

anno %>%
  inner_join(lang_df, by = "lemma") %>%
  group_by(doc_id) %>%
  summarize(percent_fr = mean(lang == "Français")) %>%
  filter(between(percent_fr, 0.4, 0.6)) %>%
  inner_join(docs, by = "doc_id") %>%
  slice_sample(n = 4) %>%
  dsst_print_text(max_chars = 10000)
```

You can see that these are reviews where someone wrote the review in one
language and then translated it into the other language.

Let's try to see how the language of each review corresponds to the reviewers.

```{r}
lang_by_user <- anno %>%
  inner_join(lang_df, by = "lemma") %>%
  group_by(doc_id) %>%
  summarize(percent_fr = mean(lang == "Français")) %>%
  mutate(lang = if_else(percent_fr > 0.75, "Français", "Mixed")) %>%
  mutate(lang = if_else(percent_fr < 0.25, "English", lang)) %>%
  inner_join(docs, by = "doc_id") %>%
  group_by(user_name) %>%
  summarize(percent_fr = mean(lang == "Français"),
            percent_mix = mean(lang == "Mixed"),
            percent_en = mean(lang == "English"))
```

To start, how does this distribution compare to the document-level one:

```{r}
lang_by_user %>%
  ggplot(aes(percent_fr)) +
    geom_histogram(bins = 50, color = "black", fill = "white")
```

And who are the reviewers that use a mixture of the two languages:

```{r}
lang_by_user %>%
  arrange(desc(percent_mix))
```

Finally, let's see how the language of the bilingual reviews from Hugo and
Mathieu  corresponds to the half of the review.

```{r}
anno %>%
  group_by(doc_id) %>%
  mutate(half = if_else(row_number() > n() * 0.5, "1st", "2nd")) %>%
  inner_join(lang_df, by = "lemma") %>%
  group_by(doc_id) %>%
  mutate(percent_fr = mean(lang == "Français")) %>%
  filter(between(percent_fr, 0.3, .7)) %>%
  group_by(half, doc_id) %>%
  summarize(percent_fr = mean(lang == "Français")) %>%
  inner_join(docs, by = "doc_id") %>%
  filter(user_name %in% c("Hugo", "Mathieu")) %>%
  group_by(half, user_name) %>%
  summarize(percent_fr = mean(percent_fr))
```

We see that Hugo almost always starts in English and follows in French. Mathieu
is more likely to start in French and switch into English, but is less
consistent.

Now, what type of business are most associated with each language? Let's check
the French reviews first:

```{r, warning=FALSE}
lang_by_user %>%
  inner_join(docs, by = "user_name") %>%
  mutate(lang = if_else(percent_fr > 0.5, "Français", "English")) %>%
  group_by(biz_category) %>%
  summarize(avg_fr = mean(lang == "Français")) %>%
  arrange(desc(avg_fr))
```

And with the English reviews?

```{r, warning=FALSE}
lang_by_user %>%
  inner_join(docs, by = "user_name") %>%
  mutate(lang = if_else(percent_fr > 0.5, "Français", "English")) %>%
  group_by(biz_category) %>%
  summarize(avg_fr = mean(lang == "Français")) %>%
  arrange(avg_fr)
```

There's a lot more that you could do here, but I thought this gave a good
example and review of how we can use unsupervised learning techniques to 
understand a rich data set.
