---
title: "07. G-score"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

```{r, message=FALSE, echo=FALSE}
source("../funs/funs.R")
options(dsst.traceit = FALSE)
```

# Load the Data

As in the previous notes, I will use the Amazon product classification task.
We will read in the `docs` and `anno` tables:

```{r, message = FALSE}
docs <- read_csv("../data/amazon_product_class.csv")
anno <- read_csv("../data/amazon_product_class_token.csv.gz")
```

These notes will briefly introduce a new measurement for associating a term
with a particular category.

## G-Score

So far, we have focused on building predictive models based on features 
composed on word counts. We've had the twin goals of identifying words that 
are strongly associated (negatively or positively) with each category and
understanding how different the categories are from one another. Can we 
easily tell them apart? Can we tell some of the apart? Can we tell some of
them apart some of the time? And so forth.

If we are only interested in the first question of what terms are associated
with each category, we can try to do this more directly measuring the strength
of the relationship between the occurance of a word and the presence of a
particular category. There are a few different such scores that are commonly
used with textual data. The one I find the best is called the G-score; it 
provides a single number that tells how strongly a single term is associated
with a given category.

We will derive the G-score more fully through the slides in class. Here we'll
just look at a function to that computes the value:

```{r}
dsst_metrics(anno, docs)
```

By default the table is ordered by category and then in ascending order by the
gscore. You'll probably want to select only the top scores for each label, which
can be done as follows:

```{r}
dsst_metrics(anno, docs) %>%
  group_by(label) %>%
  slice_head(n = 4)
```

And as with previous notes, we can modify and filter the initial dataset before
running the metrics:

```{r}
anno %>%
  mutate(lemma = if_else(upos == "PRON", tolower(token), lemma)) %>%
  mutate(lemma = if_else(lemma == "i", "I", lemma)) %>%
  filter(upos %in% c("ADJ", "ADV", "NOUN", "VERB", "PRON")) %>%
  dsst_metrics(docs) %>%
  group_by(label) %>%
  slice_head(n = 4)
```

The last column is another measurement called the chi-squared statistic. If you've
had a previous statistics course, this corresponds to the classic chi-squared test
for a 2-by-2 table. We won't use it much here, but I included it just in case you
want another measurment to compare the G-score to.
