---
title: "Notebook 14 -- Solutions"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
    css: "../css/note-style.css"
---

## Getting Started

Before running this notebook, select "Session > Restart R and Clear Output" in
the menu above to start a new R session. This will clear any old data sets and
give us a blank slate to start with.

After starting a new session, run the following code chunk to load the
libraries and data that we will be working with today.

```{r, include=FALSE, message=FALSE}
source("../funs/funs.R")
source("../funs/funs_custom.R")
```

Note that the questions today are a bit more open ended than in previous
notebooks.

## Qu'ran Surah

We are going to use an API today that gives information about the Qu'ran. The
text is broken up into 114 surah (similar to Chapters). We can get metadata 
about each of the surah by using an API defined as follows:

- protocol:   **http**
- authority:  **api.alquran.cloud**
- path:       **v1/surah**

There are no query parameters needed for this API call. In the block below,
call this API using a cache directory and save the results of the parsed JSON
data as an object named `obj`:

```{r, question-01}
url_base <- modify_url("http://api.alquran.cloud/v1/surah")
res <- dsst_cache_get(url_base, cache_dir = "cache")
obj <- content(res, type = "application/json")
```

Now, in the next block, create a tabular dataset containing as much information
as possible from the object. Use the techniques we saw today to figure out the
returned objects structure:

```{r, question-02}
# obj$data[[1]]   # note: this is a helpful object to look at

surah <- tibble(
  number = map_int(obj$data, ~ .x$number),
  name = map_chr(obj$data, ~ .x$name),
  english_name = map_chr(obj$data, ~ .x$englishName),
  english_trans = map_chr(obj$data, ~ .x$englishNameTranslation),
  ayah_count = map_int(obj$data, ~ .x$numberOfAyahs),
  rev_type = map_chr(obj$data, ~ .x$revelationType)
)

surah
```

Make a plot that shows the surah number on the x-axis and the ayah (verse)
count on the y-axis. Color the points based on the revelation type (either
Meccan or Medinian). Take note of any patterns in the way the surah are
ordered.

```{r, question-03}
surah %>%
  ggplot(aes(number, ayah_count)) +
    geom_point(aes(color = rev_type))
```

## Qu'ran Surah Text

The same service that we used above also provides the text of individual surah.
This can be done with the HTTP API using the following parameter, where # is 
replaced by the surah number.

- protocol:   **http**
- authority:  **api.alquran.cloud**
- path:       **v1/surah/#/en.asad**

In the code below, grab the data for of the 2nd surah ("The Cow") using this 
API and save the result of the parsed JSON data as `obj`:

```{r, question-04}
url_base <- modify_url("http://api.alquran.cloud/v1/surah/2/en.asad")
res <- dsst_cache_get(url_base, cache_dir = "cache")
obj <- content(res, type = "application/json")
```

Again, create a tabular dataset out of `obj`, storing as much information as
possible.

```{r, question-05}
surah_txt <- tibble(
  number = obj$data$number,
  name = obj$data$name,
  english_name = obj$data$englishName,
  english_trans = obj$data$englishNameTranslation,
  ayah_count = obj$data$numberOfAyahs,
  rev_type = obj$data$revelationType,
  edition_id = obj$data$edition$identifier,
  ayah_num = map_int(obj$data$ayahs, ~ .x$number),
  text = map_chr(obj$data$ayahs, ~ .x$text),
  juz = map_int(obj$data$ayahs, ~ .x$juz),
  manzil = map_int(obj$data$ayahs, ~ .x$manzil),
  page = map_int(obj$data$ayahs, ~ .x$page),
  ruku = map_int(obj$data$ayahs, ~ .x$ruku),
  hizb_quarter = map_int(obj$data$ayahs, ~ .x$hizbQuarter),
  sajda = map_lgl(obj$data$ayahs, ~ .x$sajda)
)

surah_txt %>%
  select(number, text)
```

There is not, yet, much we can do in terms of analysis with the text directly.
We will return to this task in the coming weeks.


